---
title: "Lab 7"
author: "C.Swick"
date: "2022-11-08"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = FALSE)
library(tidyverse)
library(here)
library(ggpubr)
library(janitor)
library(broom)
library(knitr)
```


#Load Data 
```{r}
trees_mod <- read_csv(here("data", "trees_mod.csv"))
trees <- trees_mod %>%  clean_names()
```

```{r tree, fig.align="center", fig.cap="Figure 1: scatter plot of DBH (in.) and height of the tree (ft.)."}
tree_scatter <- ggplot(trees, aes(x=height, y=girth)) + geom_point() +
  labs(x= "Height (ft.)", y= "DBH (in.)") +
  xlim(60,90)+ylim(0,25) +
  theme_bw()

tree_scatter

```


## Model 0
```{r results=TRUE}
model_0 <- lm(girth~1, data = trees)
summary(model_0)

#check the output 
mean(trees$girth)
sd(trees$girth)/sqrt(length((trees$girth)-1))

```



## Plot model_0
finding and showing mean of y 

```{r results=TRUE}
tree_scatter_0 <- ggplot(trees, aes(x=height, y=girth)) + geom_point() +
  labs(x= "Height (ft.)", y= "DBH (in.)") +
  xlim(60,90)+ylim(0,25) +
  geom_hline(yintercept = 13.25, linetype = "dashed") +
  theme_bw()

tree_scatter_0

```


## model_1 

```{r results=TRUE}
model_1 <- lm(girth~height, data = trees)
summary(model_1)

```


# Plot model_1
```{r results=TRUE}
tree_scatter_1 <- ggplot(trees, aes(x=height, y=girth)) + geom_point() +
  labs(x= "Height (ft.)", y= "DBH (in.)") +
  xlim(60,90)+ylim(0,25) +
  geom_smooth(formula = y~x, method = "lm") +
  theme_bw()

tree_scatter_1

```


## model_2 (no intercept)
we removed the intercept because it had a large standard error and r value (not enough evidence to say its different from 0), we set the intercept to 0, because tree height and dbh cant be less than 0
```{r results=TRUE}
model_2 <- lm(girth~ -1 + height, data = trees)
summary(model_2)

```


# Plot model_2
```{r results=TRUE}
tree_scatter_2 <- ggplot(trees, aes(x=height, y=girth)) + geom_point() +
  labs(x= "Height (ft.)", y= "DBH (in.)") +
  xlim(0,90)+ylim(0,25) +
  geom_smooth(formula = y~-1+x, method = "lm") +
  theme_bw()

tree_scatter_2

```



in code
> plot(model_2)
Hit <Return> to see next plot: 
Hit <Return> to see next plot: 
Hit <Return> to see next plot:



## Correlations
```{r results=TRUE}
cor.test(trees$height, trees$volume, method = "pearson")
# about 59.8% correlation
# reject null in favor of alternative, p-value is small 
```



## model_3 

```{r results=TRUE}
model_3 <- lm(girth~ volume, data = trees)
summary(model_3)
# both intercept and volume and significant different from 0 so they are both valid explanatory variables 
# for every one unit increase in volume, you add .184 inches in height 
```


explaining values of y with information from x 
~ "as a function of" y is a function of the following girth = y
R^2 amount of variation comapred to a mean only model (model_0)
  null model (mean) has lots of variability, a good linear model will have less variability 
  the R^2 is 1- variation of model divided by variation in null 
    we want a larger R^2 value 
    
## model_4

```{r results=TRUE}
model_4 <- lm(girth~ volume + height, data = trees)
summary(model_4)
# intercept and volume are significantly different from 0
#change due to height is not significantly different from 0
```


## model_5

```{r results=TRUE}
model_5 <- lm(girth~ volume * height, data = trees)
summary(model_5)
# the * tells me about significance, the intercept is not significantly different from 0
```



# Plot model_5????
```{r results=TRUE}
tree_scatter_5 <- ggplot(trees, aes(x=volume, y=girth)) + geom_point() +
  labs(x= "volume", y= "DBH (in.)") +
  xlim(0,90)+ylim(0,25) +
  geom_smooth(formula = y~x+height+height*x, method = "lm") +
  theme_bw()

tree_scatter_5

```


#making data frames and tables 
```{r results = TRUE}
out_5_tidy <- tidy(model_5) # coefficients part 
out_5_glance <- glance(model_5) # model part 
# tidy and glance in broom package 

out_5_glance
out_5_tidy
```



```{r results = TRUE}
kable(out_5_tidy, format = "markdown", digits = 3, caption = "Tests of Linear Model (model 5) coefficients")
kable(out_5_glance, format = "markdown", digits = 3, caption = "Tests of Linear Model (model 5)")
```





